---
title: "Gapminder Analysis"
author: "Stephen Turner"
date: "January 1, 2017"
output: pdf_document
---

# Introduction 
  
This is my first RMarkdown document! 
  
# Let's embed some R code 
  
Let's load the **Gapminder** data: 

```{r cache = TRUE} 
library(dplyr) 
library(readr) 
gm <- read_csv('../data/gapminder.csv') 
head(gm) 
``` 

```{r echo = FALSE, results = 'hold'} 
head(gm) 
tail(gm) 
``` 

```{r, results='asis'} 
library(knitr) 
kable(head(gm)) 
``` 

The mean life expectancy is `r mean(gm$lifeExp)` years. 

The years surveyed in this data include: `r unique(gm$year)`. 

# Session Information 

```{r} 
sessionInfo() 
``` 

# Make a figure

```{r, fig.cap='Life Exp vs GDP'} 
library(ggplot2) 
ggplot(gm, aes(gdpPercap, lifeExp)) + geom_point() 
``` 

# Make another figure

```{r, fig.cap='Life Exp vs GDP'} 
library(ggplot2) 
ggplot(gm, aes(gdpPercap, lifeExp)) +  
geom_point() +  
scale_x_log10() +  
aes(col=continent) 
```

# Options

* echo: (TRUE by default) whether to include R source code in the output file.
* results takes several possible values:
    * markup (the default) takes the result of the R evaluation and turns it into markdown that is rendered as usual.
    * hide will hide results.
    * hold will hold all the output pieces and push them to the end of a chunk. Useful if you’re running commands that result in lots of little pieces of output in the same chunk.
    * asis writes the raw results from R directly into the document. Only really useful for tables.
* include: (TRUE by default) if this is set to FALSE the R code is still evaluated, but neither the code nor the results are returned in the output document.
* fig.width, fig.height: used to control the size of graphics in the output.

```{r optionalChunkName, echo = TRUE, results = 'hide'} 
# R code here 
``` 
