---
title: "Exercise 1"
author: "Adam Lu"
date: "2018/3/12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE}
library(dplyr)
library(survival)
library(survminer)
```

Take a look at the built in `colon` dataset. If you type `?colon` it’ll ask you if you wanted help on the colon dataset from the survival package, or the colon operator. Click “Chemotherapy for Stage B/C colon cancer”, or be specific with  `?survival::colon`. This dataset has survival and recurrence information on 929 people from a clinical trial on colon cancer chemotherapy. There are two rows per person, indicated by the event type (`etype`) variable – `etype == 1` indicates that row corresponds to recurrence; `etype == 2` indicates death.

First, let’s turn the `colon` data into a tibble, then filter the data to only include the survival data, not the recurrence data. Let’s call this new object `colondeath`. The `filter()` function is in the `dplyr` library, which you can get by running  `library(dplyr)`. If you don’t have dplyr you can use the base `subset()` function instead.

```{r}
colon <- as_tibble(colon)
colon
colondeath <- filter(colon, etype == 2)
colondeath

# Or, using base subset()
# colondeath <- subset(colon, etype == 2)

head(colondeath)
```

1. Look at the help for `?colon` again. How are `sex` and `status` coded? How is this different from the lung data?

```{r eval = F}
?survival::colon
head(colon$sex)
head(colon$status)
```

2. Using `survfit(Surv(..., ...,)~..., data=colondeath)`, create a survival curve separately for males versus females. Call the resulting object `sfit`. Run a `summary()` on this object, showing time points 0, 500, 1000, 1500, and 2000. Do males or females appear to fair better over this time period?

```{r}
sfit1 <- survfit(Surv(time, status) ~ sex, data = colondeath)
summary(sfit1, times = seq(0, 2000, 500))
```

    Answer: The survival curves are about the same.

3. Using the survminer package, plot a Kaplan-Meier curve for this analysis with confidence intervals and showing the p-value. See `?ggsurvplot` for help. Is there a significant difference between males and females?

```{r}
ggsurvplot(sfit1, conf.int = TRUE, pval = TRUE,
           legend.labs = c("Female", "Male"),
           risk.table = TRUE, legend.title = "Sex")
#           title("Kaplan-Meier Curve for Colon Cancer Survival")
```

    Answer: There is no significant difference between males and females. 

4. Create Kaplan-Meier plot stratifying by:
a. The extent of differentiation (well, moderate, poor), showing the p-value.   

```{r}
sfit2 <- survfit(Surv(time, status) ~ differ, data = colondeath)
summary(sfit2, times = seq(0, 2000, 500))
ggsurvplot(sfit2, conf.int = TRUE, pval = TRUE,
           legend.labs = c("well", "moderate", "poor"),
           risk.table = TRUE, legend.title = "Differentiation")
#           title("Kaplan-Meier Curve for Colon Cancer Survival")
```

Whether or not there was detectable cancer in >=4 lymph nodes, showing the p-value and confidence bands. 

```{r}
sfit3 <- survfit(Surv(time, status) ~ node4, data = colondeath)
summary(sfit3, times = seq(0, 2000, 500))
ggsurvplot(sfit3, conf.int = TRUE, pval = TRUE,
           legend.labs = c("no", "yes"),
           risk.table = TRUE, legend.title = "At least 4 lymph nodes")
#           title("Kaplan-Meier Curve for Colon Cancer Survival")
```

Don't do this: 

```{r}
sfit4 <- survfit(Surv(time, status) ~ nodes, data = colondeath)
ggsurvplot(sfit4)
```
