---
title: "Exercise 3"
author: "Adam Lu"
date: "2018/3/12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE}
library(dplyr)
library(survival)
library(survminer)
library(RTCGA)
library(RTCGA.clinical)
```

The **“KIPAN”** cohort (in `KIPAN.clinical`) is the pan-kidney cohort, consisting of KICH (chromaphobe renal cell carcinoma), KIRC (renal clear cell carcinoma), and KIRP (papillary cell carcinoma). The `KIPAN.clinical` has `KICH.clinical`, `KIRC.clinical`, and `KIPR.clinical` all combined.

1. Using `survivalTCGA()`, create a new object called `clinkid` using the `KIPAN.clinical` cohort. For the columns to extract, get both the disease code and the patient’s gender (`extract.cols=c("admin.disease_code", "patient.gender")`). The first few rows will look like this.

```{r}
clinkid <- survivalTCGA(KIPAN.clinical, 
                        extract.cols = c("admin.disease_code",
                                         "patient.gender"))
head(clinkid)
```

2. The xtabs() command will produce tables of counts for categorical variables. Use this command to examine how many samples you have for each kidney sample type, separately by sex.

```{r}
xtabs(~ admin.disease_code + patient.gender, data = clinkid)
```


3. Run a Cox PH regression on the cancer type and gender. What’s the effect of gender? Is it significant? How does survival differ by each type? Which has the worst prognosis?

```{r}
coxph(Surv(times, patient.vital_status) ~ 
              admin.disease_code + patient.gender, data = clinkid)
```

    Answer: Male had slightly better survival but the effect was not significant. Papillary cell carcinoma had the worst prognosis. 

4. Create survival curves for each different subtype. 
    a. Produce a Kaplan-Meier plot. 

```{r}
sfit <- 
    survfit(Surv(times, patient.vital_status) ~ admin.disease_code,
            data = clinkid)

ggsurvplot(sfit, conf.int = TRUE, pval = TRUE)
```

    b. Show survival tables each year for the first 5 years.

```{r}
summary(sfit, times=seq(0, 365*5, 365))
```
