---
title: "Exercise 3"
author: "Adam Lu"
date: "2018/03/10"
output: html_document
---

First, import libraries

```{r, message = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
```

Next, import the [National Health and Nutrition Examination Survey data](https://www.cdc.gov/nchs/nhanes/index.htm)

```{r, message = FALSE}
nh <- read_csv("../data/nhanes.csv")
nh
```

Change all character data types to the "factor" data type

```{r, message = FALSE}
nh <- nh %>%
    mutate_if(is.character, as.factor)
nh
```

Filter the data to include adults only

```{r}
nha <- nh %>% filter(Age >= 18)
nha
```

1. What’s the relationship between diabetes and participating in rigorous physical activity or sports?

* Create a contingency table with Diabetes status in rows and physical activity status in columns.
```{r}
xt <- xtabs(~ Diabetes + PhysActive, data = nha)
xt
```
    
* Display that table with margins.

```{r}
addmargins(xt)
```
    
* Show the proportions of diabetics and nondiabetics, separately, who are physically 
active or not.

```{r}
prop.table(xt, margin = 1)
```
    
* Is this relationship significant?

```{r}
# Use the Pearson's Chi-squared test with Yates' continuity correction
chisq.test(xt)

# Fisher's Exact Test for Count Data
# https://en.wikipedia.org/wiki/Fisher%27s_exact_test
fisher.test(xt)
```
    
* Create a mosaic plot to visualize the relationship.

```{r}
mosaicplot(xt, main = "Diabetics are less physically active")
```
    
2. Model the same association in a logistic regression framework to assess the risk of diabetes using physical activity as a predictor.

* Fit a model with just physical activity as a predictor, and display a model summary.

```{r}
fit <- glm(Diabetes ~ PhysActive, data = nha, family = "binomial")
summary(fit)
```
    
* Add gender to the model, and show a summary.

```{r}
fit <- glm(Diabetes ~ PhysActive + Gender, data = nha, family = "binomial")
summary(fit)
```
    
* Continue adding weight and age to the model. What happens to the gender association?

```{r}
fit <- glm(Diabetes ~ PhysActive + Gender + Weight + Age, data = nha, family = "binomial")
summary(fit)
```
    
* Continue and add income to the model. What happens to the original association with physical activity?

```{r}
fit <- glm(Diabetes ~ PhysActive + Gender + Weight + Age + Income, data = nha, family = "binomial")
summary(fit)
```
    
    
