---
title: "Exercise 4"
author: "Adam Lu"
date: "2018/03/10"
output: html_document
---

1. You’re doing a gene expression experiment. What’s your power to detect a 2-fold change in a gene with a standard deviation of 0.7, given 3 samples? (Note - fold change is usually given on the log2 scale, so a 2-fold change would be a delta of 1. That is, if the fold change is 2x, then log2(2)=1, and you should use 1 in the calculation, not 2).

```{r}
power.t.test(n = 3, delta = 1, sd = 0.7)$power
```

2. How many samples would you need to have 80% power to detect this effect?

```{r}
power.t.test(power = 0.8, delta = 1, sd = 0.7)$n
```

3. You’re doing a population genome-wide association study (GWAS) looking at the effect of a SNP on disease X. Disease X has a baseline prevalence of 5% in the population, but you suspect the SNP might increase the risk of disease X by 10% (this is typical for SNP effects on common, complex diseases). Calculate the number of samples do you need to have 80% power to detect this effect, given that you want a genome-wide statistical significance of p<5×10−8 to account for multiple testing.1 (Hint, you can expressed 5×10−8 in R using 5e-8 instead of .00000005).

```{r}
power.prop.test(power = 0.8, p1 = 0.05, p2 = 0.055, 
                sig.level = 5e-8, alternative = "two.sided")$n
```

